---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

github_sha_link <- function(sha) {
  glue::glue("[{sha}](https://github.com/gadenbuie/js4shiny-frappeCharts/commit/{sha})")
}
```


# Building a Shiny Input

In this project, 
we're going to create a typing speed app 
using a custom Shiny input.
The app will give users typing prompts,
monitor their typing speed,
and use a Frappe Chart line chart 
to show their speed over time as they type.

## Setup a folder for our app inside the frappeCharts package

`r github_sha_link("113340074c3af9c2cdf46cd7787829d4ec56bfcf")`

Create the directory `inst/shiny-input-app` and add `app.R` and `typing.js`.

```{r}
usethis::use_directory("inst/shiny-input-app")
file.create("inst/shiny-input-app/app.R")
file.create("inst/shiny-input-app/typing.js")
```

## Create a basic Shiny app with a typing area

`r github_sha_link("ff3a962f7e6d16a75ebdb620aac0fdfc9949086e")`

We'll start with typical Shiny inputs.

```r
library(shiny)

ui <- fluidPage(
  textAreaInput("typing", "Type here..."),
  verbatimTextOutput("debug")
)

server <- function(input, output, session) {
  output$debug <- renderPrint(input$typing)
}

shinyApp(ui, server)
```

Run this app and type in the box.

## Create our own typingSpeedInput()

Use Shiny's `textAreaInput()` to get the template 
for our own `typingSpeedInput()`

```{r, eval=TRUE}
shiny_text_input <- shiny::textAreaInput(
  "INPUT", "LABEL", placeholder = "PLACEHOLDER"
)
cat(format(shiny_text_input))
```


```{r}
typingSpeedInput <- function(inputId, label, placeholder = NULL) {
  .label <- label
  htmltools::withTags(
    div(
      class = "form-group typing-speed",
      label(class = "control-label", `for` = inputId, .label),
      textarea(id = inputId, class = "form-control", placeholder = placeholder)
    )
  )
}
```

Two points:

1. Notice that I used `htmltools::withTags()`,
   which makes it easier to write multiple tags at once.
   But it has the downside of masking the `label` argument of `typingSpeedInput()`.
   Hence, the first line `.label <- label`.

1. I added `.typing-speed` to our parent container so that we can
   find or style our custom input.
   
Replace the `textAreaInput()` with our new `typingSpeedInput()` and run the app.
It works the same!
Wait, why?

```{r}
ui <- fluidPage(
  # textAreaInput("typing", "Type here..."),
  typingSpeedInput("typing", "Type here..."),
  verbatimTextOutput("debug")
)
```

