---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

github_sha_link <- function(sha) {
  glue::glue("[{sha}](https://github.com/gadenbuie/js4shiny-frappeCharts/commit/{sha})")
}
```


# Building a Shiny Input

In this project, 
we're going to create a typing speed app 
using a custom Shiny input.
The app will give users typing prompts,
monitor their typing speed,
and use a Frappe Chart line chart 
to show their speed over time as they type.

## Setup a folder for our app inside the frappeCharts package

`r github_sha_link("2188781657884a743c2c4f94dc1fa7755a7de261")`

Create the directory `inst/shiny-input-app` and add `app.R` and `typing.js`.

```{r}
usethis::use_directory("inst/shiny-input-app")
file.create("inst/shiny-input-app/app.R")
file.create("inst/shiny-input-app/typing.js")
```

## Create a basic Shiny app with a typing area

We'll start with typical Shiny inputs.

```r
library(shiny)

ui <- fluidPage(
  textAreaInput("typing", "Type here..."),
  verbatimTextOutput("debug")
)

server <- function(input, output, session) {
  output$debug <- renderPrint(input$typing)
}

shinyApp(ui, server)
```

Run this app and type in the box.
