---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

github_sha_link <- function(sha) {
  glue::glue("[{sha}](https://github.com/gadenbuie/js4shiny-frappeCharts/commit/{sha})")
}

get_last_sha <- function() {
  commit <- git2r::commits()[[1]]
  message(commit$summary)
  message(commit$sha)
  clipr::write_clip(glue::glue(
    "`r github_sha_link(\"{commit$sha}\")`"
  ))
}
```


# Building a Shiny Input

In this project, 
we're going to create a typing speed app 
using a custom Shiny input.
The app will give users typing prompts,
monitor their typing speed,
and use a Frappe Chart line chart 
to show their speed over time as they type.

## Setup a folder for our app inside the frappeCharts package

`r github_sha_link("113340074c3af9c2cdf46cd7787829d4ec56bfcf")`

Create the directory `inst/shiny-input-app` and add `app.R` and `typing.js`.

```{r}
usethis::use_directory("inst/shiny-input-app")
file.create("inst/shiny-input-app/app.R")
file.create("inst/shiny-input-app/typing.js")
```

## Create a basic Shiny app with a typing area

`r github_sha_link("ff3a962f7e6d16a75ebdb620aac0fdfc9949086e")`

We'll start with typical Shiny inputs.

```r
library(shiny)

ui <- fluidPage(
  textAreaInput("typing", "Type here..."),
  verbatimTextOutput("debug")
)

server <- function(input, output, session) {
  output$debug <- renderPrint(input$typing)
}

shinyApp(ui, server)
```

Run this app and type in the box.

## Create our own typingSpeedInput()

`r github_sha_link("b7108029b58635652ce87f3e1ea9a2c5a6232020")`

Use Shiny's `textAreaInput()` to get the template 
for our own `typingSpeedInput()`

```{r, eval=TRUE}
shiny_text_input <- shiny::textAreaInput(
  "INPUT", "LABEL", placeholder = "PLACEHOLDER"
)
cat(format(shiny_text_input))
```


```{r}
typingSpeedInput <- function(inputId, label, placeholder = NULL) {
  .label <- label
  htmltools::withTags(
    div(
      class = "form-group typing-speed",
      label(class = "control-label", `for` = inputId, .label),
      textarea(id = inputId, class = "form-control", placeholder = placeholder)
    )
  )
}
```

Two points:

1. Notice that I used `htmltools::withTags()`,
   which makes it easier to write multiple tags at once.
   But it has the downside of masking the `label` argument of `typingSpeedInput()`.
   Hence, the first line `.label <- label`.

1. I added `.typing-speed` to our parent container so that we can
   find or style our custom input.
   
Replace the `textAreaInput()` with our new `typingSpeedInput()` and run the app.
It works the same!
Wait, why?

```{r}
ui <- fluidPage(
  # textAreaInput("typing", "Type here..."),
  typingSpeedInput("typing", "Type here..."),
  verbatimTextOutput("debug")
)
```

## Start creating an input binding for `typingSpeedInput()`

`r github_sha_link("02adb896c50a97ebe7f9a7fef2a6f00488f9418d")`

Now we can open `typing.js` and create a Shiny input binding.

If you used `js4shiny::snippets_install()`,
you have a `ShinyInputBinding` snippet that provides a template for you.
Or you can copy the chunk below.

<details><summary>Shiny Input Binding Template</summary>

```js
// Ref: https://shiny.rstudio.com/articles/building-inputs.html
// Ref: https://github.com/rstudio/shiny/blob/master/srcjs/input_binding.js

const bindingName = new Shiny.InputBinding();

$.extend(bindingName, {
  find: function(scope) {
    // Specify the selector that identifies your input. `scope` is a general
    // parent of your input elements. This function should return the nodes of
    // ALL of the inputs that are inside `scope`. These elements should all
    // have IDs that are used as the inputId on the server side.
    return scope.querySelectorAll("inputBindingSelector");
  },
  getValue: function(el) {
    // For a particular input, this function is given the element containing
    // your input. In this function, find or construct the value that will be
    // returned to Shiny. The ID of `el` is used for the inputId.

    // e.g: return el.value
    return 'FIXME';
  },
  setValue: function(el, value) {
    // This method is used for restoring the bookmarked state of your input
    // and allows you to set the input's state without triggering reactivity.
    // Basically, reverses .getValue()

    // e.g.; el.value = value
    console.error('bindingName.setValue() is not yet defined');
  },
  receiveMessage: function(el, data) {
    // Given the input's container and data, update the input
    // and its elements to reflect the given data.
    // The messages are sent from R/Shiny via
    // R> session$sendInputMessage(inputId, data)
    console.error('bindingName.receiveMessage() is not yet defined');

    // If you want the update to trigger reactivity, trigger a subscribed event
    $(el).trigger("change")
  },
  subscribe: function(el, callback) {
    // Listen to events on your input element. The following block listens to
    // the change event, but you might want to listen to another event.
    // Repeat the block for each event type you want to subscribe to.

    $(el).on("change.bindingName", function(e) {
      // Use callback() or callback(true).
      // If using callback(true) the rate policy applies,
      // for example if you need to throttle or debounce
      // the values being sent back to the server.
      callback();
    });
  },
  getRatePolicy: function() {
    return {
      policy: 'debounce', // 'debounce', 'throttle' or 'direct' (default)
      delay: 100 // milliseconds for debounce or throttle
    };
  },
  unsubscribe: function(el) {
    $(el).off(".bindingName");
  }
});

Shiny.inputBindings.register(bindingName, 'pkgName.bindingName');
```

</details>

If you use the snippet,
it automatically walks you through 
the first pass of parts that need to be changed.
If you copied the template,
you need to find and replace all:

- `bindingName`: the name of the JavaScript object with the specifics of your
  Shiny input. This name isn't user facing, but helps Shiny keep track of which
  the inputs in an app.
  
  We'll call ours `typingSpeed`
  
- `inputBindingSelector`: this is the CSS selector that can be used to find
  the HTML element of your input. The element you find here is the element in
  your input's HTML that has the `id` with the input's `inputId`.
  
  In our case, we want the `textarea` element 
  that's a descendent of `.typing-speed`, so we use
  
  ```css
  .typing-speed textarea
  ```
  
- `change`: This is the event that will be listened to by Shiny to know when
  the input has updated. For our typing speed input, we're going to listen to
  the `keyup` event.
  
- `pkgName`: if you're writing this input as part of a package, use your package
  name. It's not critical; this just provides some namespacing in case there's
  another package that impelements an in put with the same `bindingName`.
  
  Here we use `js4shiny`.
