---
output: github_document
---

# Building an HTML Widget: Frappe Charts

[htmlwidget]: https://www.htmlwidgets.org/
[frappe-charts]: https://frappe.io/charts
[rmd-book-htmlwidget]: https://bookdown.org/yihui/rmarkdown/html-widgets.html

This repository walks through the building of an [htmlwidget]
around the JavaScript library [Frappe Charts][frappe-charts].

Because there are many files and moving pieces involved in this process,
I've created a git repository that walks through the changes
at each step of the process.
With my notes about each step,
I've included the SHA linked to the updates made during the step.
If you're viewing this on GitHub,
those SHA hashes should be converted to links 
that will take you to a summary of which files changed at each step.

This introduction focuses on the mechanics of htmlwidgets.
For a much more detailed summary,
the [HTML Widgets][rmd-book-htmlwidget] chapter of the R Markdown book 
is an excellent introduction.


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

github_sha_link <- function(sha) {
  glue::glue("[changelog: {substr(sha, 1, 6)}](https://github.com/gadenbuie/js4shiny-frappeCharts/commit/{sha})")
}
```

## Setup R Package

* `r github_sha_link("a3f6fdd986d2b98323b5be43e323df4f6a19f1f3")`

Create a package for this HTML widget.
We're not going to publish this, so you can call it whatever you want

```{r create-package}
usethis::create_package("frappeCharts")
```

Add a dev script for notes

```{r dev}
dir.create("dev")
file.create("dev/dev.R")
rstudioapi::navigateToFile("dev/dev.R")
```


### Add the R package dependencies for an htmlwidget package

```{r r-deps}
usethis::use_package("htmlwidgets")
usethis::use_package("htmltools")
usethis::use_package("jsonlite")
usethis::use_package("shiny")
usethis::use_package("yaml")
```

## Setup npm package

* `r github_sha_link("256f0ca112b2685608f9a17a4fb4e35d279c9830")`

Same process again, but this time for npm.

```bash
npm init

# or

npm init -y
```

Open `package.json` and take a look

```json
{
  "name": "frappecharts",
  "version": "0.0.1",
  "description": "",
  "main": "index.js",
  "scripts": {
    "test": "echo \"Error: no test specified\" && exit 1"
  },
  "author": "",
  "license": "MIT"
}
```

From Frappe Charts [docs#installation](https://frappe.io/charts/docs#installation):

```bash
npm install frappe-charts
```

We now have a dependency in `package.json` and there's a `package-lock.json` file.

```json
"dependencies": {
  "frappe-charts": "^1.3.0"
}
```

## Ignore node_modules but add package-lock

There's also a `node_modules/` folder with `frappe-charts/` inside.
Add `node_modules` to `.Rbuildignore` and `.gitignore`.
(BTW, you can and are supposed to commit `package-lock.json`.)

```{r ignore-node-module}
usethis::use_build_ignore("node_modules")
usethis::use_build_ignore("package.json")
usethis::use_build_ignore("package-lock.json")
usethis::use_git_ignore("node_modules")
```

## Scaffold the HTML widget

* `r github_sha_link("38bac2c65cf54816525076690310008e62ab99a1")`

```{r htmlwidgets-scaffold}
htmlwidgets::scaffoldWidget("frappeChart")
```

This adds files in `inst/htmlwidgets`

```
inst
└── htmlwidgets
    ├── frappeChart.js    #<< R <-> JS code
    └── frappeChart.yaml  #<< list of dependencies
```

and creates a file `R/frappeChart.R` with the functions

- `frappeChart()`
- `frappeChartOutput()` (for shiny)
- `renderFrappeChart()` (for shiny)

## Use `npm` to get our dependencies in the right place

* `r github_sha_link("7abf0224345a67217c4a476f04eafe581f0ecec0")`

`htmlwidgets` load dependencies in a way that's exactly the same as using a
`<script>` tag in the HTML `<head>`.
Look at the 
[documentation on Frappe Charts](https://frappe.io/charts/docs#installation)
and decide which file we should use.

Here's the block from their docs

```
<script src="https://cdn.jsdelivr.net/npm/frappe-charts@1.2.4/dist/frappe-charts.min.iife.js"></script>
<!-- or -->
<script src="https://unpkg.com/frappe-charts@1.2.4/dist/frappe-charts.min.iife.js"></script>
```

We need to get our dependecy into a subfolder of `inst/htmlwidgets`.
Convention is `inst/htmlwidgets/lib/<dependency_name>`.
Rather than creating the directoy and copying over, etc.,
we can have an `npm` build script do this for us.

To avoid issues with mac/windows,
we'll add a dev dependency on [`cpy-cli`](https://github.com/sindresorhus/cpy-cli).
Dev dependencies are node modules
that are used to build a package,
rather than required for the package to work.

```bash
npm install cpy-cli --save-dev
```

Then we create the folder `frappe-charts` under `inst/htmlwidgets/lib`
that will hold the Frappe Charts JavaScript dependency.
(If the library included other required files,
we would move these too.)

```{r create-lib-dir}
dir.create("inst/htmlwidets/lib/frappe-charts", recursive = TRUE)
```

And then edit `package.json` to add a copy task.
You can define scripts that are runnable with `npm run <script-name>`.
For small build tasks, this is an easy to implement build solution.

```
"scripts": {
  "copy-js": "cpy 'node_modules/frappe-charts/dist/frappe-charts.min.iife*' inst/htmlwidgets/lib/frappe-charts/",
  "build": "npm run copy-js"
}
```

Notice that running `npm run build` will also call `npm run copy-js`.
If we had more build tasks related to our JavaScript dependencies,
like linting or testing,
we could add them as separate scripts
and have them run in the build process with 
`npm run <task-1> && npm run <task-2>`, etc.

## Create a demo html_document_plain()

* `r github_sha_link("036d454f80d6036fc1ba35db92161fd19c053635")`

```{r create-demo-html}
dir.create("dev/demo")
js4shiny::js4shiny_rmd(path = "dev/demo/demo.Rmd")
```

Use the example in the [Frappe Charts Docs](https://frappe.io/charts/docs).

```{r}
tagList(
  div(id = "chart"),
  htmltools::htmlDependency(
    name = "frappe-charts",
    version = "1.3.0",
    package = "frappeCharts",
    src = "htmlwidgets/lib/frappe-charts",
    script = "frappe-charts.min.iife.js",
    all_files = TRUE
  )
)
```

And copy the JS into a javascript chunk.

`r emo::ji("warning")` The dependencies won't be found until you build/install.

```{r build-install}
devtools::document()
devtools::install()
```

If you get a path not found error

```
Error: path for html_dependency not found: inst/htmlwidgets/lib/frappe-charts
```

it's most likely because

```
src = "inst/htmlwidgets/lib/frappe-charts"
```

should be relative to `inst`.

## Replace the example data with another data set and example

* `r github_sha_link("8fd703a08b021b8466171b83506f5fb0bf92f2ac")`

The first demo mixes chart types and we don't want to do that.
Use the example from
[Basic Chart](https://frappe.io/charts/docs/basic/basic_chart#adding-more-datasets).

```js
const data = {
    labels: ["Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"],
    datasets: [
      { name: "R", values: [18, 40, 30, 35, 8, 52, 17, -4] },
      { name: "Python", values: [30, 50, -10, 15, 18, 32, 27, 14] }
    ]
}
```

Then re-create this data in an R chunk (`r github_sha_link("881c12ffdbdaa017863c918f61fa6208400d6130")`:

```{r data-in-r}
data <- list(
  labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
  datasets = list(
    list(name = "R", values = c(18, 40, 30, 35, 8, 52, 17, -4)),
    list(name = "Python", values = c(30, 50, -10, 15, 18, 32, 27, 14))
  )
)
```

To get the data out of R and make it available in the document,
`htmlwidgets` embeds the data in a `<script type="application/json">...</script>`
element in the page.
Embed the data from the R chunk in a `<script>` tag with an ID 
so that we can find it later.

```{r embed-r-data-in-script}
tags$script(
  id = "data",
  type = "application/json",
  htmlwidgets:::toJSON(data)
)
```

Change to `js4shiny::html_document_js()` so that we can see the `console.log()`
from JavaScript just like R code.
And then find the `<script>` tag and get it's `.textContent`.

```{js find-r-data-script-2}
let rData = document.getElementById('data')
rData.textContent
```

Use `JSON.parse()` to turn the data into a JS object
and replace the data used in the chart (`r github_sha_link("7201e436e72ebddee271cbf7c02a733ac81a5d86")`.

```{js find-r-data-script-3}
let rData = document.getElementById('data')
rData = JSON.parse(rData.textContent)
```

Switch between `data` and `rData` and it should be the same!

Change the values of the data in the R side to be random
so that each re-run gives a new plot.

~~Delete the `data` in the JS side.~~
Comment out the `data` on the JS side (but we'll want to see the structure later).

## Augment data to set options for the chart

* `r github_sha_link("3e1d9bee03fdf621f5dc5ec46e0e92f603ebe219")`

Embed `data` in another list `opts` that will carry additional options, 
such as `title`, `type` and `colors`.

Parse the embedded `<script>` and pass the whole object to `frappe.Chart()`.

Change the colors to 

- `#466683` (dark blue)
- `#44bc96` (green)
- `#d33f49` (red)
- `#993d70` (purple)

## Learn about other options for line charts

* `r github_sha_link("340d516ee4c7788e4f7e5089c4957ee9ffd1333e")`

Read <https://frappe.io/charts/docs/basic/trends_regions>
and add and test additional line options.

Goal: shaded area chart with lines only.

Make the `labels` one week and repeat 4 times.
Generate `runif(7 * 4)` random numbers.

```r
rep(c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"), 4)
```

Find and implement an option to reduce the number of labels on the x-axis.

## Turn on dots again and make navigable

* `r github_sha_link("93d4c74f4b30a819b5c22fd7cc8ff238fc62f572")`

```{r}
opts <- list(
  title = "My AwesomeR Chart",
  type = "bar",
  height = 250,
  colors = c("#466683", "#44bc96"),
  data = data,
  axisOptions = list(xIsSeries = TRUE),
  isNavigable = TRUE
)
```

## Add a real data source

* `r github_sha_link("7a988739e3b5ff0572f4c16ce5110f52936550c3")`

This is where you decide how much work you want to do on the R side
and how much work should be done on the JavaScript side.
One thing is clear though,
R users should not be expected 
to construct the nested list data structure
just to use your HTML widget.
We are used to data.frames and tibbles,
so these should be supported out of the box!

To practice our JavaScript skills, 
I've chosen to do most of the work in the browser.
We'll just ask our users to give us rectangular data,
where the first column will provide the x axis labels 
and the remaining columns are series.

Sidenote: As you can tell, there's a lot of validation that should happen,
but that part is not as much fun so we're going to pretend
our users will always give us perfectly formatted data.
If you do take on an htmlwidget package project,
_don't skimp on this step_.
Having a friendly R API will have huge impact on the use of your widget.

We'll use the `babynames` package for our demo dataset, 
pick two names completely at random to compare.

```{r babynames}
library(dplyr)
library(babynames)

data <-
  babynames %>% 
  filter(
    name %in% c("Ruth", "August"),
    year >= 1980
  ) %>% 
  group_by(year, name) %>% 
  summarize(n = sum(n)) %>% 
  ungroup() %>% 
  pivot_wider(year, name, values_from = n)
```

At this point the chart won't work,
but you can use the browser dev console 
to find the right steps to reformat the data into the expected format.

We'll make the **strong** assumption that the tibble in R
should always be formatted with the columns

1. `labels`
2. first series...
3. second series...
4. etc.

* `repl_example('reformat-r2js-data')`

<details><summary>Answer</summary>

```js
const chartData = {labels: [], datasets: []}

// Get keys of data, assume that first entry is for labels, the rest are data
let labelColumn = Object.keys(x.data)[0]
let columns = Object.keys(x.data).slice(1)

// First column in x.data is the labels
chartData.labels = x.data[labelColumn]

// Create an appropriate object for each column, reformat data and add to chartData
columns.forEach(function(col) {
  chartData.datasets.push({name: col, values: x.data[col]})
})

x.data = chartData
```

</details>


## This is basically what `htmlwidgets` does, just inside a framework

* `r github_sha_link("a0614d9699aefc0eda82b3e368b48370be0ae9ba")`

We now have all of the pieces of an `htmlwidget`,
it's just a bit less coordinated.

1. `htmlwidgets` gives us a slightly nicer way of specifying dependencies
   in `inst/htmlwidgets/frappeChart.yaml`. We'll have to update that file.
   
2. When we started we added a `div(id = "chart")`. 
   It would be annoying to have to make sure that each `id` is always unique.
   `htmlwidgets` will add this `div` for us and give each one a unique id.
   We won't have to write any code for this, it just happens.
   
3. We'll write an R function that will take input data and options and format
   it into a list, like the `opts` we've been using.
   Then we hand the data to `htmlwidgets` and it embeds it in a `<script>` tag
   for us.
   It will also find that data automatically and make it available on the JS side.
   
4. Finally, we wrote some code in JavaScript to initialize the chart.
   In the same way, we'll write some code in `inst/htmlwidgets/frappeChart.js`
   which is where we'll reformat the data and options passed from the R world
   by htmlwidgets. We also need to instantiate the chart object. For advanced
   usage, this is also where we'll put code that would let us update the
   widget in place without having to re-render the whole chart.
   
To create the htmlwidget,
we're going to work through each of these pieces
and put them in the right places.

# Make it an htmlwidget

## Declare dependencies

* `r github_sha_link("969fd962edf0be8f98ffff1823f8e08960ffb31a")`

**FILE:** `inst/htmlwidgets/frappeChart.yaml`

Take the `htmltools::htmlDependency()` and 
turn it into `inst/htmlwidgets/frappeChart.yaml`.

```{r}
rstudioapi::navigateToFile("inst/htmlwidgets/frappeChart.yaml")
```

Note: keep `htmlwidgets` in `src`!

## Write the R function

* `r github_sha_link("cbc25a8f7bbf7e522f369f5fec7c2517ba768656")`

**FILE:** `R/frappeChart.r`

Add appropriate arguments to `frappeChart()`.

* [title](https://frappe.io/charts/docs/reference/configuration#title)
* [type](https://frappe.io/charts/docs/reference/configuration#type)
* [colors](https://frappe.io/charts/docs/reference/configuration#colors)?
* [is_navigable](https://frappe.io/charts/docs/reference/configuration#isnavigable)

Structure the arguments into `x` and pass `...` for the "extra bits".

Rebuild the package,
then create a new R markdown document:
`js4shiny::js4shiny_doc()`.
 
Move the code loading `dplyr`, `tidyr`, `babynames`
and formatting the data.
Then call `frappeCharts::frappeChart()`.

Render and open dev tools in the browser to see that it "works".
Meaning that the data and dependencies are included,
but the chart won't.
Point out the random ID.
Then go back and change it so we can find the element better.

## Write JavaScript binding

* `r github_sha_link("6f141c4341a2c4f8615df81887e7927d2e765f11")`

**FILE:** `inst/htmlwidgets/frappeChart.js`

The final step is to move the Javascript we wrote before into the js binding.

* Just put in `console.log(x)`, rebuild, rerender
* Verify that this `x` looks the same as our `opts` from before
* Copy all of the JS we wrote to reconfigure the data into the widget
* Use `el` instead of `#chart`
* Rebuild, rerender
* it works!
* Try adding other options

### Writing JavaScript in R

* `r github_sha_link("8d442e3c842154adbae87dab5e9289cbb1333187")`

The [tooltips](https://frappe.io/charts/docs/basic/annotations#tooltips)
can be formatted using the `tooltipOptions` property:

```
tooltipOptions: {
    formatTooltipX: d => (d + '').toUpperCase(),
    formatTooltipY: d => d + ' pts',
}
```

To write this in R (add to `widget_demo.R`)

```r
tooltipOptions = list(
  formatTooltipX = htmlwidgets::JS("d => 'Year: ' + d"),
  formatTooltipY = htmlwidgets::JS("d => d + ' babies'")
)
```

## Shiny comes for free!

* `r github_sha_link("739d5945010d5e46ab3f9847fd412beb0766805d")`

Create a basic Shiny app with

1. Slider input to pick number of values (1:26 letters)
1. A new data button that generates new data of same dimension
1. The data are reactive, `x = letters[1:n]`, `y = runif(n)`
1. Use `frappeCharts::frappeChartOutput()` linked to `frappeCharts::renderFrappeChart()`
    - bar plot
    - fix `tooltipOptions` to turn the `runif()` into a percent.
    
`dev/shiny/app.R`

Make a mistake in the spelling for `formatTooltipY`
and demo how hard it is for the end user to track down what's wrong.
This points to how important it is to do the validation on the R side
or to do the extra work to make the R API friendly.

It's also a good place to demo debug strategies for Shiny and regular widgets.
Open the app in an external window,
show the dev console,
find the frappeCharts binding
and add a breakpoint.
Then reload and show how you an use the dev console there to figure things out.

## Better data updates

Frappe Charts, 
like many JS libraries,
includes a method for updating the widget
without having to redraw the whole chart/plot/viz/etc.

In Frappe Charts, the 
[full data update](https://frappe.io/charts/docs/update_state/modify_data#updating-full-data)
method is

```js
chart.update(data)
```

where `data` is the `data` part of the initial options object.

Let's make this work...

### Refactor the JS-side data processing code

`r github_sha_link("b19e33af8fdca579a8578bcd7a39c6d1e43fb32c")`

- Create a `prepareChartData()` function from the code we wrote for
  `renderValue()`. The goal is that this will let us use the function in 
  multiple places.
  
#### Make `chart` generally available

`r github_sha_link("d114592668ca63f06f593f4f247432eec218894b")`
      
Make the created `chart` object available outside `renderValue()`

#### Expose the context inside the factory function to the world (and yourself)

`r github_sha_link("f0a3bf9fd5e60cda9b2b7ace004f360c36bf6610")`

- bind the factory function context to `el` as `widget`
- Demo this by opening a rendered widget and showing `widget` as attached to the div

#### Expose `chart` with a `chart()` method

`r github_sha_link("f0a3bf9fd5e60cda9b2b7ace004f360c36bf6610")`

Add a chart method to the widget object so that we can get to the current chart.

1. Demo by finding widget div and running
   
   ```
   let c = $0.widget.chart()
   c.addDataPoint(2017, [2500, 1500])
   ```
1. Now, if nothing else, the `chart` object is accessible
   so others can use or extend it.
   
#### Create an update method that takes new data and updates an existing chart
   
`r github_sha_link("5da4b68b5f60d8e6ee17cc8c4a009121539a2653")`

Demo with `app.R`

```js
let el = document.getElementById('chart')
el.widget.update({x: ['A', 'B', 'C', 'D'], Frequency: [1, 2, 3, 4]})
```

Try with various values. You can increase the number of data points
but you can't add or change the series.

#### Add a custom message handler that dependes on `HTMLWidgets.shinyMode`

`r github_sha_link("5da4b68b5f60d8e6ee17cc8c4a009121539a2653")`

```js
// after factory function
if (HTMLWidgets.shinyMode) {
  Shiny.addCustomMessageHandler('frappeCharts:update', function({id, data}) {
    let el = document.getElementById(id)
    el.widget.update(data)
  })
}
```

Restructure the app code so that the chart initializes with flat data (0.5).
Use `session$sendCustomMessage` to trigger the update.

Note that the JS function above takes `id` and `data` using destructuring.
It's easy to write `function(id, data)` but this won't work because
the handler can only take one argument.

Demo the app, now updates are fast!

#### Write a user-friendly wrapper around `sendCustomMessage` called `updateFrappeChart()` 

`r github_sha_link("4706d89183aaa9a3721599ef13c6f7af4955808b")`

#### Now add an event listener to send chart navigation back to Shiny 

`r github_sha_link("0b4f7ea16f378ec5a53d81260c8f9056fabbcaba")`
    
Attach the event listener during `renderValue()` and watch for the `data-select` event.
Use the `el.id` to create a new id, like `el.id + '_selected'`.
Send back `index` and `values` from the event.

Add `verbatimTextOutput('selected')` to show `input$chart_selected`.

#### Return better values

`r github_sha_link("e7fe0e1d87977823e6a040434b33e6d5cdf8eac1")`

You would probably want to do some work for the user and return more meaningful values.
We'll probably just copy and paste this during the workshop, 
but here's a potential method.

This function basically reverses the chart processing and 
and returns a list that should be a dataframe.

```js
if (HTMLWidgets.shinyMode && x.isNavigable) {
el.addEventListener('data-select', function(ev) {
  let {index, values} = ev
  let chart = el.widget.chart()
  let label = chart.data.labels[index]
  let names = chart.data.datasets.map(d => d.name)
  let data = values.reduce(function(acc, v, idx) {
    acc[names[idx]] = v
    return acc
  }, {})
  data[labelsName] = label
  Shiny.setInputValue(el.id + '_selected', data)
})
}
```

#### Process the returned data for the user in Shiny

`r github_sha_link("000de60582f277e29983f6c5803de112ca1ade99")`

But now in Shiny it needs to go from a list to a data.frame.
To do this we use `shiny::registerInputHandler()` in R and
give the input event a type: `inputId_selected:frappeCharts-selected`.

```r
.onLoad <- function(libname, pkgname) {
  shiny::registerInputHandler(
    type = "frappeCharts-selected",
    fun = function(value, session, inputName) {
      as.data.frame(value, stringsAsFactors = FALSE)
    }
  )
}
```
