```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Setup R Package

Create a package for this HTML widget.
We're not going to publish this, so you can call it whatever you want

```{r create-package}
usethis::create_package("frappeCharts")
```

Add a dev script for notes

```{r dev}
dir.create("dev")
file.create("dev/dev.R")
rstudioapi::navigateToFile("dev/dev.R")
```

## Setup npm package

Same process again, but this time for npm.

```bash
npm init

# or

npm init -y
```

Open `package.json` and take a look

```json
{
  "name": "frappecharts",
  "version": "0.0.1",
  "description": "",
  "main": "index.js",
  "scripts": {
    "test": "echo \"Error: no test specified\" && exit 1"
  },
  "author": "",
  "license": "MIT"
}
```

From Frappe Charts [docs#installation](https://frappe.io/charts/docs#installation):

```bash
npm install frappe-charts
```

We now have a dependency in `package.json` and there's a `package-lock.json` file.

```json
"dependencies": {
  "frappe-charts": "^1.3.0"
}
```

## Ignore node_modules but add package-lock

There's also a `node_modules/` folder with `frappe-charts/` inside.
Add `node_modules` to `.Rbuildignore` and `.gitignore`.
(BTW, you can and are supposed to commit `package-lock.json`.)

```{r ignore-node-module}
usethis::use_build_ignore("node_modules")
usethis::use_build_ignore("package.json")
usethis::use_build_ignore("package-lock.json")
usethis::use_git_ignore("node_modules")
```

## Scaffold the HTML widget

```{r htmlwidgets-scaffold}
htmlwidgets::scaffoldWidget("frappeChart")
```

This adds files in `inst/htmlwidgets`

```
inst
└── htmlwidgets
    ├── frappeChart.js    #<< R <-> JS code
    └── frappeChart.yaml  #<< list of dependencies
```

and creates a file `R/frappeChart.R` with the functions

- `frappeChart()`
- `frappeChartOutput()` (for shiny)
- `renderFrappeChart()` (for shiny)

## Use `npm` to get our dependencies in the right place

`htmlwidgets` load dependencies in a way that's exactly the same as using a
`<script>` tag in the HTML `<head>`.
Look at the documentation on Frappe Charts and decide which file we should use.

Here's the block from their docs

```
<script src="https://cdn.jsdelivr.net/npm/frappe-charts@1.2.4/dist/frappe-charts.min.iife.js"></script>
<!-- or -->
<script src="https://unpkg.com/frappe-charts@1.2.4/dist/frappe-charts.min.iife.js"></script>
```

We need to get our dependecy into a subfolder of `inst/htmlwidgets`.
Convention is `inst/htmlwidgets/lib/<dependency_name>`.
Rather than creating the directoy and copying over, etc.,
we can have an `npm` build script do this for us.

To avoid issues with mac/windows,
we'll add a dev dependency on [`cpy-cli`](https://github.com/sindresorhus/cpy-cli)

```bash
npm install cpy-cli --save-dev
```

and

```{r create-lib-dir}
dir.create("inst/htmlwidets/lib/frappe-charts", recursive = TRUE)
```

and then edit `package.json` to add copy tasks

```
"scripts": {
  "copy-js": "cpy 'node_modules/frappe-charts/dist/frappe-charts.min.iife*' inst/htmlwidgets/lib/frappe-charts/",
  "build": "npm run copy-js"
}
```

## Create a demo html_document_plain()

```{r create-demo-html}
dir.create("dev/demo")
js4shiny::js4shiny_rmd(path = "dev/demo/demo.Rmd")
```

Use the example in the [Frappe Charts Docs](https://frappe.io/charts/docs).

```{r}
tagList(
  div(id = "chart"),
  htmltools::htmlDependency(
    name = "frappe-charts",
    version = "1.3.0",
    package = "frappeCharts",
    src = "htmlwidgets/lib/frappe-charts",
    script = "frappe-charts.min.iife.js",
    all_files = TRUE
  )
)
```

And copy the JS into a javascript chunk.

`r emo::ji("warning")` The dependencies won't be found until you build/install.

```{r build-install}
devtools::document()
devtools::install()
```

If you get a path not found error

```
Error: path for html_dependency not found: inst/htmlwidgets/lib/frappe-charts
```

it's most likely because

```
src = "inst/htmlwidgets/lib/frappe-charts"
```

## Replace the example data with another data set and example

The first demo mixes chart types and we don't want to do that.
Use the example from
[Basic Chart](https://frappe.io/charts/docs/basic/basic_chart#adding-more-datasets).

```js
const data = {
    labels: ["Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"],
    datasets: [
      { name: "R", values: [18, 40, 30, 35, 8, 52, 17, -4] },
      { name: "Python", values: [30, 50, -10, 15, 18, 32, 27, 14] }
    ]
}
```

Then re-create this data in an R chunk:

```{r data-in-r}
data <- list(
  labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
  datasets = list(
    list(name = "R", values = c(18, 40, 30, 35, 8, 52, 17, -4)),
    list(name = "Python", values = c(30, 50, -10, 15, 18, 32, 27, 14))
  )
)
```

To get the data out of R and make it available in the document,
`htmlwidgets` embeds the data in a `<script type="application/json">...</script>`
element in the page.
Embed the data from the R chunk in a `<script>` tag with an ID 
so that we can find it later.

```{r embed-r-data-in-script}
tags$script(
  id = "data",
  type = "application/json",
  htmlwidgets:::toJSON(data)
)
```

Change to `js4shiny::html_document_js()` so that we can see the `console.log()`
from JavaScript just like R code.
And then find the `<script>` tag and get it's `.textContent`.

```{js find-r-data-script}
let rData = document.getElementById('data')
rData.textContent
```

Use `JSON.parse()` to turn the data into a JS object
and replace the data used in the chart.

```{js find-r-data-script}
let rData = document.getElementById('data')
rData = JSON.parse(rData.textContent)
```

Switch between `data` and `rData` and it should be the same!

Change the values of the data in the R side to be random
so that each re-run gives a new plot.

Delete the `data` in the JS side.

